Open Preprints in Ecology \& Evolution

### 2012-09-14 ms.tex

Preprint servers such as arXiv are common in mathematics and physics,
but still a minority of papers in ecology and evolution use
them. Preprint servers allow authors to make their manuscripts
publicly available before, or at the same time as, submitting them to
journals for traditional peer-review. ...  The idea became popular
with arXiv, and many physicists start their day with a look at the new
papers in ArXiv \cite{gin11}

We will highlight advantages for both scientists and publishers.

The first and most often discussed advantage is speed \ref{fig:map}.

@@

%% @Tim: how can it improve the review process?

@@

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will
% not give your MS to anyone!
Prepublication reviews by a small network of colleagues is an
important part of the scientific process, which is attested by the
fact that nearly all published papers acknowledge comments by people
not listed as co-authors.  Preprint servers simply offer a way to
extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small
networks of scientists exchanging ideas.  Paul Ginsparg created
arXiv.org in part for democratic reasons: he wanted everyone from
students in small universities to Ivy-League professors to have access
to the most recent scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg's revolutionary idea was simply to use the power of the
internet for preprints, not just for the end product, so the process
can be open from A to Z, instead of being just open at the end of the
process.
% TP Since when is the publishing buisness an open process? :-)

@@

% A short paragraph to conclude.

Responding to the rumour that they refused manuscripts submitted to arXiv,
Nature responded that ``Nature never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.
Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original
community-building purposes.  Preprint servers also facilitate
communication between disciplines, bridging cultural as well as
geographic divides.
% JA: perhaps something less advertising-like should go here:
The advantages are clear and the costs are low.
% BUY NOW!

>>>>

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. ...

@@

@@

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will
% not give your MS to anyone!
Prepublication reviews by a small network of colleagues is an
important part of the scientific process, which is attested by the
fact that nearly all published papers acknowledge comments by people
not listed as co-authors.  Preprint servers simply offer a way to
extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small
networks of scientists exchanging ideas.  Paul Ginsparg created
arXiv.org in part for democratic reasons: he wanted everyone from
students in small universities to Ivy-League professors to have access
to the most recent scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg's revolutionary idea was simply to use the power of the
internet for preprints, not just for the end product, so the process
can be open from A to Z, instead of being just open at the end of the
process.
% TP Since when is the publishing buisness an open process? :-)
% PhDP: Wishful thinking :) Corrected...

Preprint servers also establish priority in a fair way. Since some manuscripts
will spend much more time in the review process, public preprints servers offer
a fairer way to establish intellectual priority by making the work available
when done. Surprisingly, there is perception in biology that public preprints
make it easier to steal ideas, while mathematicians and physicists have embraced
arXiv in part to establish intellectual priority in a fair way \cite{cal12}.

@@

Responding to the rumour that they refused manuscripts submitted to arXiv,
Nature responded that ``Nature never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.

Open preprints server offer a great opportunity for open science, especially if
the community embrace the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work. These initiatives are
important to fully exploit the potential of open preprints servers.

Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original
community-building purposes.  Preprint servers also facilitate
communication between disciplines, bridging cultural as well as
geographic divides.
% JA: perhaps something less advertising-like should go here:
The advantages are clear and the costs are low.
% BUY NOW!
